---
title: "Exploratory Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("rate_box_plot.R")
library(dplyr)
library(ggplot2)
obesity_data <- read.csv("Obesity_GDP_PanelData.csv")
```

# Distribution of Variables 
### Obesity and Poverty Rates
```{r, echo=FALSE}
boxplot(rates$obesity_rate, rates$poverty_rate,
    main = "Distribution of Rates",
    names = c("Obesity Rate", "Poverty Rate"),
    col = rainbow(2),
    ylab = "Percentages"
)
```

The boxplots above display the distribution of the obesity rates and poverty rates across all fifty states. It is clear that the distribution of obesity rates is higher than that of poverty rates. The average obesity rate across all states is higher than that of poverty rates. The mean obesity rate is `r signif(obesity_avg, 3)`% which is `r signif(obesity_avg - poverty_avg, 3)`% higher than the poverty mean of `r signif(poverty_avg, 3)`%. In addition, the median obesity rate of `r signif(obesity_median, 3)`% is `r signif(obesity_median - poverty_median, 3)`% higher than the poverty rate median of `r signif(obesity_median, 3)`%. This shows that, on average, a higher percent of people across the United States are obese then are in poverty. Both rates have an inner quartile range of under 5, meaning that 50% of the rates for each states are less than a 5% difference of each other. The inner quartile range for obesity rates is `r signif(obesity_Q3, 3)` and the inner quartile range for poverty rates is `r signif(poverty_Q3, 3)`. The distribution of both rates can also be analized in terms of standard deviation. The standard deviation for obesity rate is `r signif(obesity_sd, 3)` and the standard deviation for poverty rate is `r signif(poverty_sd, 3)`.


# Relationships Between Variables
## Effects Of Wealth and Obesity
``` {r, echo = FALSE}
ggplot(obesity_data, 
                                    aes(x = Average.Income,
                                        y = Adult.Obesity.100,
                                        color = Year)) +
  geom_point() +
  xlab("Average Income") +
  ylab("Percentage of Adult Obesity By State And Year") +
  ggtitle("Income vs Obesity Rate")
```
<br>
Unsuprisingly, we see a correlation between income and obesity. In general as the average 
income increases, we see the obesity rate decreases as well. This isn't a surprise due 
to the issue that less wealthy people are more likely to live in food deserts and have less access
to purchasing healthier foods. Another issue that can influence this correlation is that 
junk food are typically cheaper than healthier foods. We see this with fast food restaurants
and how they are able to offer very cheap meals as compared to a standard restaurant. 

``` {r, echo = FALSE}
ggplot(obesity_data,
                                  aes(x = Poverty.Rate.100,
                                      y = Adult.Obesity.100,
                                      color = Year)) +
  geom_point() +
  xlab("Poverty Rate Percentage in State") + 
  ylab("Percentage of Adult Obesity By State And Year") +
  ggtitle("Poverty Rate Vs. Obesity Rate")
```
<br>
We continue to see the trend with wealth and obesity rates as well when we compare poverty
rates and obestiy rates. In this graph, we see that as poverty rates rises, so does the 
obesity rate. This falls in line with our conclusions with average income because income 
is the determining factor if someone lives in poverty or not.

``` {r, echo = FALSE}
ggplot(obesity_data,
                         aes(x = Real.GDP,
                             y = Adult.Obesity.100,
                             color = Year)) +
  geom_point() +
  xlab("State GDP") + 
  ylab("Percentage of Adult Obesity By State And Year") +
  ggtitle("State GDP Vs. Obesity Rate")
```
<br>
Finally, comparing a state's GDP to its obesity rate, we see that there isn't a strong
correlation between the two. This is highly surprising because the previous two charts
display a fairly strong correlation between wealth and obesity while as we see in this chart,
a state's GDP typically doesn't affect the obesity rates of its state. Although, we can see
that the wealthier states have a relatively low obesity rate, the poorer states have a large variation
of obesity rates. This draws into question if there are other factors, such as region, that can cause such
a large disparity among the poor states.

Another interesting trend we see in all three charts is that the years, which are color 
coded on the graphs are distrubuted quite evenly. We do not see clusters of obesity and income rates for certain years which suggests that there is not a signicant change in income or obesity levels as time goes on. However, our data is limited from 2014-2017 so further research will be needed to better determine how obesity levels change with time. 